---
title: "Unit 2: Rtutorial"
author: "Jennifer Krzak"
format: html
editor: visual
---

## Week 6: Working with Data

In the previous R tutorial, we started to work on science classroom dataset. We applied the data intensive research steps to explore our data and investigate the relationship between students' grades and time-spent.

Let's remember which libraries and functions we used!

#### [**Your Turn:**]{.underline}

Please write down one or two sentence to explain why and how we use the following libraries and functions.

1.  **tidyverse:**

2.  **skimr:**

3.  **ggplot:**

4.  **read_csv():**

5.  **view():**

6.  **glimpse():**

7.  **head():**

8.  **tail():**

9.  **select():**

10. **filter():**

11. **arrange():**

12. **desc():**

13. **geom_histogram():**

14. **geom_point():**

### Load the Tidyverse Package

Let's start our R code along by loading the tidyverse package.

```{r}
library(tidyverse)
```

### Load the online science class data

Now, load the online science class data from the data folder and assign your data to a new object.

```{r}
sci_classes<-read_csv("Data/sci-online-classes.csv")
```

You loaded the data, now what should we do?

```{r}
sci_classes
```

#### [**Your Turn:**]{.underline}

Examine the contents of sci_classes in your console. You should type the object name to the console and check that.

**Question:** Is your object a tibble? How do you know?

**Your response here:**

**Hint:** Check the output in the console.

### Check your data with different functions

You can check your data with different functions. Let's remember how we use different functions to check our data.

```{r}
view(sci_classes)
head(sci_classes)
tail(sci_classes)
glimpse(sci_classes)
```

### **Isolating Data with dplyr**

We will use select() function to select the following columns from our data.

-   student_id

-   subject

-   semester

-   FinalGradeCEMS

-   After selecting these columns, assign that to a new object with a name of "student_grade".

```{r}
student_grade<-sci_classes |>
  select(student_id,
         subject,
         semester,
         FinalGradeCEMS)
student_grade
```

#### [**Your Turn:**]{.underline}

Examine students' grades, what did you realize about it?

**Your response here:**

**Hint:** Check the missing data.

### **Specific select**

Now, we will make a specific selection.

-   Select all columns except subject and semester.

-   Assign to a new object with a different name.

-   Examine your data frame.

```{r}
new_object<-select(sci_classes,-c(subject,semester))
```

Checking the data frame:

```         
```

#### [**Your Turn:**]{.underline}

-   Select all columns except student_id and FinalGradeCEMS.

-   Assign to a new object with a different name.

-   Examine your data frame.

```{r}
your_turn_new_object<-select(sci_classes,-c(student_id,FinalGradeCEMS))
```

### **Specific select**

-   Select only the columns that start with Time

-   Assign to a new object with a different name.

-   Use view() function to examine your data frame.

```{r}

time_object<-select(sci_classes,starts_with("time"))
```

#### [**Your Turn:**]{.underline}

-   Select only the columns that ends with "r"

-   Assign to a new object with a different name.

-   Use view() function to examine your data frame.

```{r}
your_turn_time_object<-select(sci_classes,ends_with("r"))
```

### Filter Function

-   Filter the sci_classes data frame for just males.

-   Assign to a new object.

-   Use the head() function to examine your data frame.

```{r}
male_sort<-sci_classes|>
  filter(Gender=="M")

head(male_sort)
```

#### [**Your Turn:**]{.underline}

-   Filter the sci_classes data frame for just females.

-   Assign to a new object.

-   Use the tail() function to examine your data frame.

```{r}
female_sort<-sci_classes|>
  filter(Gender=="F")

tail(female_sort)
```

Let's try filter function with two arguments now.

-   Filter the sci_classes data frame for students whose

-   percentage_earned is greater than 0.8

-   Assign to a new object.

-   Use the tail() function to examine your data frame. 

```{r}

bio_student<-filter(sci_classes,percentage_earned>0.8 & subject=="BioA")

tail(bio_student)
```

[**Your Turn:**]{.underline}

Filter the sci_classes data frame for students whose

-   percentage_earned is smaller or equal to 0.6

-   Assign to a new object.

-   Use the head() function to examine your data frame. 

```{r}

bio_student2<-filter(sci_classes,percentage_earned<0.6 & subject=="FrScA")

head(bio_student2)
```

Let's use filter () function for the missing data.

-   Filter the sci_classes data frame so rows with 

-   NA for points earned are removed.

-   Assign to a new object.

-   Use the glimpse() function to examine your data frame.

```{r}
clean_data<-sci_classes|>
  filter(!is.na(Points_Earned))

glimpse(clean_data)
```

Filter the sci_classes data for the following subjects:

-   BioA

-   PhysA

-   OcnA

-   Assign to a new object with a different name.

-   Use the summary() function to examine your data frame.

```{r}
my_classes<-filter(sci_classes, subject %in% c("BioA", "PhysA", "OcnA"))

my_classes_2<-sci_classes|>
  filter(subject %in% c("BioA", "PhysA", "OcnA"))
```

### Arrange () Function

Let's recall how we were using the arrange () function for our dataset.

-   Arrange sci_classes by subject subject then 

-   percentage_earned in descending order.

-   Assign to a new object.

-   Use the str() function to examine your data frame.

```{r}
order_class<-sci_classes|>
  arrange(subject, desc(percentage_earned))

str(order_class)
```

## %\>% Pipe Operator

Using sci_classes data and the %\>% pipe operator:

-   Select subject, section, time spent in hours and final course grade.

-   Filter for students in OcnA courses with grades greater than or equal to 60.

-   Arrange grades by section in descending order.

-   Assign to a new object.

Examine the contents using a method of your choosing.

```{r}
idek_fr<-sci_classes %>%
  select(subject, section, TimeSpent_hours, FinalGradeCEMS) %>%
  filter(subject=="OcnA" & FinalGradeCEMS>60) %>%
  arrange(desc(section))

view(idek_fr)
```

### Deriving info with dplyr

We will practice summarise () and group_by () functions now.

### Summarise () Function

Using sci_classes data and the summarise() function:

-   Get a distinct count of course ids.

-   Use the %\>% operator

```{r}
sci_classes %>%
  summarise(courses=n_distinct(course_id))
```

-   Get a distinct count of course ids.

-   Use the \|\> operator

```{r}
sci_classes |>
  summarise(courses=n_distinct(course_id))
```

### Group_by () Function

Using the sci_classes data and the pipe operator.

-   Filter final grades to remove NAs. \>!is.na

-   Group your data by subject and gender.

-   Summarise your data to calculate the following stats:

-   total number of students

-   mean final grade

-   mean time spent in the course

-   Assign to a new object

-   Examine the contents using a method of your choosing.

```{r}
apple_cider<-sci_classes %>%
  filter(!is.na(FinalGradeCEMS)) %>%
  group_by(subject,Gender) 
  summarise(total=sum(student_id), 
            grade=mean(FinalGradeCEMS), 
            time=mean(TimeSpent))
```

### Mutate () Function

Replace the dashed lines in the following code to;

-   Create a new variable called score that is the product of percentage earned and 100

-   Create a faceted scatter plot with hours spent in the course on the x-axis, score on the y-axis, and point colored by gender.

-   Include an alpha value to your graph.

```{r eval=FALSE}
sci_classes %>%
  mutate(score = percentage_earned * 100) %>%
  ggplot() +
  geom_point(mapping=aes(x = TimeSpent_hours,
                           y = Points_Earned,
                           color = Gender)) +
  facet_wrap(~subject)

```

## Final Step:

You are almost done, all you need to is to render your file and publish it in one of the following platform.

#### [**Your Turn:**]{.underline}

**Render File:** For now, we will wrap up this work by converting our work into a webpage that can be used to communicate your learning and demonstrate some of your new R skills. To do so, you will need to "render" your document by clicking the Render button in the menu bar at that the top of this file. This will do two things; it will:

1.  check through all your code for any errors; and,

2.  create a file in your directory that you can use to share you work through [Posit Cloud](https://posit.cloud/learn/guide#publish-from-cloud), [RPubs](https://rpubs.com "RPubs") , [GitHub Pages](https://pages.github.com "GitHub Pages"), [Quarto Pub](https://quartopub.com "Quarto Pub"), or other methods.

3.  Submit your link to the Blackboard!

Now that you've finished your Rtutorial study, scroll back to the very top of this Quarto Document and change the author: "YOUR NAME HERE" to your actual name surrounded by quotation marks like so: author: "Dr. Cansu Tatar".

#### 
